---
title: "P8105 Homework Assignment 1"
author: "Kylie Wheelock Riley"
date: "9/20/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


# Problem 1

## Dataframe
```{r prob_1_dataframe}
###Setting the seed, so problem 1 random normal distribution stays the same
set.seed(1)

prob1_df = tibble(
  normal_var = rnorm(8),
  logical_var = normal_var > 0,
  charac_var = c("R", "is", "the", "best", "coding", "language", "ever", "!"),  
  factor_var = factor(c("negative", "low", "negative", "high", "low", "negative", "low", "low"))
)

```


## Testing means of the variables in the dataframe
```{r testing_means, warning = FALSE}
mean(pull(prob1_df, normal_var))
mean(pull(prob1_df, logical_var))
mean(pull(prob1_df, charac_var))
mean(pull(prob1_df, factor_var))
```

Both the numeric and logical variables generated a mean while the character and factor variables returned NA for the mean.

## Using as.numeric
```{r converting_to_numeric, eval = FALSE}
as.numeric(pull(prob1_df, logical_var))
as.numeric(pull(prob1_df, charac_var))
as.numeric(pull(prob1_df, factor_var))

```

After using the as.numeric function, both the logical and factor variables were converted to numeric. The logical variable contained True and False statements statements that were converted to "1" and "0" respectively. The  factor variable contained the levels Negative, Low, and High which were converted to "3", "2", and "1", respectively. The character variable was not converted into a numeric. There is no level or binary structure that can be assigned to the character variable, so it was not converted. This demonstrates why the character variable was not able to generate a mean, and why the logical variable did not have a mean originally, but would could after being converted to a numeric variable.

## Converting vectors and multiplying
```{r converting_chunk, warning = FALSE}
##convert the logical vector to numeric, and multiply the random sample by the result
as.numeric(pull(prob1_df, logical_var))*(pull(prob1_df, normal_var))

##convert the logical vector to a factor, and multiply the random sample by the result
as.factor(pull(prob1_df, logical_var))*(pull(prob1_df, normal_var)) 

##convert the logical vector to a factor and then convert the result to numeric, and multiply the random sample by the result
as.numeric(as.factor(pull(prob1_df, logical_var)))*(pull(prob1_df, normal_var))

```

Converting a logical vector into a factor does not allow you to multiply it by a numeric, it needs to be converted from logical to factor and then to numeric. 



# Problem 2

## Dataframe
```{r}
prob2_df = tibble(
  x = rnorm(500),
  y = rnorm(500),
  logical_var2 = (x + y > 1),
  numeric_var2 = as.numeric(logical_var2),
  factor_var2 = as.factor(logical_var2),
)

```

```{r in line code chunk, include = FALSE}
mean_samp = mean(pull(prob2_df, x))
median_samp = median(pull(prob2_df, x))
sd_samp = sd(pull(prob2_df, x))


prob_logical = 1 - pnorm(1, mean = 0.02, sd = 1.01)
percent_log = 100*prob_logical
```

The dataset above has `r nrow(prob2_df)` rows and `r ncol(prob2_df)` columns. Variable x has a **mean** of `r round(mean_samp, digits = 2)`, a **median** of `r round(median_samp, digits = 2)`, and a **standard deviation** of `r round(sd_samp, digits = 2)`. The proportion of cases in which x + y > 1 is `r round(percent_log, digits = 2)`%.

## Scatterplots

```{r yx_scatter, echo = FALSE}
ggplot(prob2_df, aes(x = x, y = y, color = logical_var2)) + geom_point()

ggsave("logicalvarplot.pdf")

ggplot(prob2_df, aes(x = x, y = y, color = numeric_var2)) + geom_point()

ggplot(prob2_df, aes(x = x, y = y, color = factor_var2)) + geom_point()

```

The scatter plots using the logical and factor variables both contain 2 colors, one for true and one for false. The numeric variable has a gradient as the color scale. 
